---
title: "Exercise 5"
author: "Mingcong Pan"
date: "11/5/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions

* Create a function for the mean, median, and standard deviation.
```{r results = FALSE, message=FALSE, warning=FALSE}
mean <- function(x) {
 sum(x) / (length(x) - 1)
}

median <- function(x) {
  n <- length(x)
  s <- sort(x)
  ifelse(n%%2==1,s[(n+1)/2],mean(s[n/2+0:1]))
}

sd <- function(x) {
  sum((x - mean(x))^2) / (length(x) - 1)
}
```

* Create a function that finds the mean and excludes the lowest and highest value.
```{r results = FALSE, message=FALSE, warning=FALSE}
mean2 <- function(x) {
  mean(x[x > min(x) & x < max(x)])
}
```

* Apply the functions to the Christianity variables.
```{r results = FALSE, message=FALSE, warning=FALSE}
df <- read.csv("national.csv")

lapply(df[4:9], mean)
lapply(df[4:9], median)
lapply(df[4:9], sd)
lapply(df[4:9], mean2)
```

* Write a function that lists all the unique years with more than 300,000 Christians in total.
```{r results = FALSE, message=FALSE, warning=FALSE}

years <- function(df) {
  unique(df$year[df$christianity_all > 300000])
}

```

# Loops/apply

* Write a loop to find how many variables there are per observation.
```{r results = FALSE, message=FALSE, warning=FALSE, results = "hide"}
sum = 0
for(i in 1:ncol(df)){
  sum = sum + 1
}

```

* Write a loop to find the mean number of Protestant Christians in each country (i.e., the state column). Then use an apply family function to do the same.
```{r results = FALSE, message=FALSE, warning=FALSE, results = "hide"}

tapply(df$christianity_all, df$state, mean)

```

* Check the column type for each variable.
```{r results = FALSE, message=FALSE, warning=FALSE, results = "hide"}

for(i in 1:ncol(df)){
  typeof(df[,i])
}
```


